A Survey of Wearable Device Risk Perceptions

### 2015-01-05 wearables_survey.tex
finished the related work section.

\subsection{Ubiquitous Sensing}
Due to new technological advances, we are rapidly moving towards automated capture and ubiquitous data access~\cite{abowd2000charting}, whether it be through smartphones, smarthomes, networked devices, or, more recently, wearables. Wearables have new capabilities that allow capture of unique data about the user and his/her surroundings. As information capture becomes pervasive and subtle, people are naturally concerned about privacy and security~\cite{palen2003unpacking}. Privacy in ubiquitous systems can be achieved by adhering to principles such as notice, choice, and consent~\cite{langheinrich2001privacy}.

~\cite{Egelman2015} {\color{red} Serge: talk about this; you wrote it and it's not out yet.}
~\cite{bellotti1993design}: framework for designing for privacy in ubicomp.
~\cite{hong2004privacy}: privacy risk model for ubicomp. 
~\cite{jiang2002approximate}: social privacy model for ubicomp.

\subsection{Smartphones and Wearables Concerns}
Designs to improve end-user security need to be grounded in user research to better understand end-user concerns. Felt \etal were among the first to examine the security concerns of smartphone users using a large-scale online survey~\cite{Felt}. Their survey asked 3,115 smartphone users about 99 risks associated with various smartphone privileges. Participants were asked how upset they would be if a certain action had occurred without their permission. Participants rated each situation on a Likert scale ranging from ``indifferent (1)'' to ``very upset (5).'' Our work builds upon this study, by aiming to evaluate the types of concerns that users have surrounding future wearable device capabilities, and then to compare those concerns to concerns surrounding more established technologies.

@@

~\cite{chin2012measuring}: understanding user attitudes toward security and privacy for smartphones. 
~\cite{felt2012android}: if attention to, understand, and act on permission information during installation.
~\cite{laurier2001people}: Why people say where they are during mobile phone calls
~\cite{palen2000going}: behavior and practice of new mobile phone users

@@

Fischhoff \etal performed a seminal technology risk perception experiment in which they studied how people perceived the risks and benefits of 30 common activities and technologies~\cite{Fischhoff}. In their study, participants were prompted to separately rate the risks and benefits of 30 activities and technologies. They were primed to think about all people affected by the technology, and to think about long-term vs. short-term risks and benefits.. The participants rated these technologies with respect to each other on a numerical scale, being instructed to rate the least risky or least beneficial technology a 10 and scaling the ratings linearly (i.e. a technology with risk rating 20 would be considered twice as risky compared to a technology with a risk rating of 10). Using this mythology, they were able to categorize different technologies based on whether they were high-risk/high-benefit, low-risk/low-benefit, and so forth.

~\cite{acquisti2005privacy} Privacy and decision making: ppl lack enough info to make privacy-sensitive decisions; even if, likely to trade off long-term privacy for short-term benefits.
~\cite{jensen2005privacy}: Self-reports versus observed behavior
~\cite{spiekermann2001privacy}: Privacy Preferences Versus Actual Behavior

@@

%(REDO, Should I even say this?) I messed up that motorcycle question. I wish I actually had a calibration point for high risk high benefit for the Fischoff technology assessment questions. But well, none of the new technologies fit that description so we didn't really need it critically.

@@

%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Research Ethics} 
We received advance approval from University of California, Berkeley's Institutional Review Board to perform this user study. Survey data was collected in an anonymous manner. Although the focus group was not conducted in an anonymous manner, participants were not asked to offer any confidential or sensitive information.

\section{Acknowledgments}
The authors would like to thank The National Science Foundation's Graduate Research Fellowships Program (NSF GRFP), and the Intel Science and Technology Center for Secure Computing (also known as  Secure Computing Research for Users' Benefit, SCRUB), for the support of this research. Inspiration for this work emerged out of fruitful discussions at the Berkeley Laboratory for Usable and Experimental Security (BLUES) meetings. Many thanks go out to the many generous colleagues who have provided feedback and encouragement.

@@

% The following two commands are all you need in the initial runs of your .tex file to produce the bib
%  and remember to run: latex bibtex latex latex to resolve all references

>>>>

\subsection{Ubiquitous Sensing}
Due to new technological advances, we are rapidly moving towards automated capture and ubiquitous data access~\cite{abowd2000charting}, whether it be through smartphones, smarthomes, or miscellaneous networked devices. Many have worked to discover how privacy will be preserved in world of nearing omni-present sensors. Examples of such efforts include frameworks to design for privacy environments ~\cite{bellotti1993design} and evaluating ubiquitous computing applications ~\cite{scholtz2004toward}. There are also various models for understanding privacy with respect to ubiquitous computing, including a risk-based model ~\cite{hong2004privacy} and a social-based model ~\cite{jiang2002approximate}.

Wearables are the new frontier of ubiquitous computing, with new capabilities can capture unique data about the user and his/her surroundings. As information capture becomes pervasive and subtle, people are naturally concerned about privacy and security~\cite{palen2003unpacking}. Privacy in ubiquitous systems can be achieved by adhering to principles such as notice, choice, and consent~\cite{langheinrich2001privacy}. An example of research which helps such principles include work by Egelman \etal , which evaluates various indicator designs to better communicate data capture on ubiquitous computing platforms.

\subsection{Smartphones and Wearables Concerns}
Only efforts to improve end-user security which are grounded in user research directly address end-user concerns.The current state of smartphone security and privacy can be attributed to studies conducted to understand user attitudes toward security and privacy for smartphones ~\cite{chin2012measuring}, observe behaviors of new smartphone users ~\cite{palen2000going}, investigate user understanding of permissions ~\cite{felt2012android}.

Among such studies, Felt \etal were among the first to examine the security concerns of smartphone users using a large-scale online survey~\cite{Felt}. Their survey asked 3,115 smartphone users about 99 risks associated with various smartphone privileges. Participants were asked how upset they would be if a certain action had occurred without their permission. Participants rated each situation on a Likert scale ranging from ``indifferent (1)'' to ``very upset (5).'' Our work builds upon this study, by aiming to evaluate the types of concerns that users have surrounding future wearable device capabilities, and then to compare those concerns to concerns surrounding more established technologies.

@@

%~\cite{laurier2001people}: Why people say where they are during mobile phone calls

@@

Fischhoff \etal performed a seminal technology risk perception experiment in which they studied how people perceived the risks and benefits of 30 common activities and technologies~\cite{Fischhoff}. In their study, participants were prompted to separately rate the risks and benefits of 30 activities and technologies. They were primed to think about all people affected by the technology, and to think about long-term vs. short-term risks and benefits. The participants rated these technologies with respect to each other on a numerical scale, being instructed to rate the least risky or least beneficial technology a 10 and scaling the ratings linearly (i.e. a technology with risk rating 20 would be considered twice as risky compared to a technology with a risk rating of 10). Using this mythology, they were able to categorize different technologies based on whether they were high-risk/high-benefit, low-risk/low-benefit, and so forth.

While user perceptions are insightful, it has been shown that people generally lack enough information to make privacy-sensitive decisions; even if they do, they are likely to trade off long-term privacy for short-term benefits ~\cite{acquisti2005privacy}. Actual behavior deviates from self-reported behaviors ~\cite{jensen2005privacy} and privacy preferences ~\cite{spiekermann2001privacy}.

@@

@@

%COMMENTED OUT FOR ANONYMITY!!

%\section{Research Ethics} 
%We received advance approval from University of California, Berkeley's Institutional Review Board to perform this user study. Survey data was collected in an anonymous manner. Although the focus group was not conducted in an anonymous manner, participants were not asked to offer any confidential or sensitive information.

%\section{Acknowledgments}
%The authors would like to thank The National Science Foundation's Graduate Research Fellowships Program (NSF GRFP), and the Intel Science and Technology Center for Secure Computing (also known as  Secure Computing Research for Users' Benefit, SCRUB), for the support of this research. Inspiration for this work emerged out of fruitful discussions at the Berkeley Laboratory for Usable and Experimental Security (BLUES) meetings. Many thanks go out to the many generous colleagues who have provided feedback and encouragement.

@@
